---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  fig.retina = 2
)
```

# systemaGlobalis

systemaGlobalis is an R package that provides quick access to the complete [Gapminder data repository](https://github.com/open-numbers/ddf--gapminder--systema_globalis).

## Installation

You can install the development version of systemaGlobalis from [GitHub](https://CRAN.R-project.org) with:

``` r
devtools::install_github("mvuorre/systemaGlobalis")
```

It is not currently available on CRAN (nor is it likely to be in the near future.)

## Description

systemaGlobalis is a minimal R package interface to the complete [Systema Globalis](https://github.com/open-numbers/ddf--gapminder--systema_globalis#systema-globalis-sg) data compilation gathered by Gapminder.

```{r packages, results = 'hide', message = FALSE}
library(systemaGlobalis)
library(tidyverse)
theme_set(theme_linedraw() + theme(panel.grid = element_blank()))
```

It containts the following data frames (tibbles):

- `datapoints`
  - For each country and year, 532 variables
  - The main data frame you'll want to work with.
  
```{r}
dim(datapoints)
datapoints[1:5, 1:5]  # First five rows and columns
```
  
- `countries`
  - Additional information, such as naming schemes, regions, etc., about the countries included in `datapoints`
  
```{r}
countries
```

- `concepts`
  - Descriptions of each of the variables in `datapoints`, including sources, tags, etc.

```{r}
concepts
```

That's it. There are no functions, simply clean datasets that are easily used in analyses. 

## Examples

Let's first look at a selection of variables containing the word "income":

```{r}
datapoints %>% 
  select(contains("income")) %>% 
  names(.)
```

```{r}
datapoints %>% 
  ggplot(aes(time, income_per_person_long_series)) +
  geom_line(aes(group = geo)) +
  scale_x_continuous(limits = c(1900, 2018))
```

To find out more about this particular variable, look at the `concepts` data frame

```{r}
concepts %>% 
  filter(str_detect(concept, "income")) %>%
  select(1:2) %>% 
  slice(1:4) %>% 
  kable()
```

We can use the `countries` data frame to get more information about the countries

```{r}
datapoints %>% 
  left_join(rename(countries, geo = country)) %>% 
  filter(!is.na(world_4region)) %>% 
  ggplot(aes(time, income_per_person_long_series)) +
  geom_line(aes(group = geo)) +
  scale_x_continuous(limits = c(1800, 2018)) +
  facet_wrap("world_4region")
```

And, of course, animations! Here's an animated version of the figure on page 26 of Factfulness:

```{r animation, cache = TRUE}
library(gganimate)
p <- datapoints %>% 
  select(geo, 
         time, 
         child = children_per_woman_total_fertility,
         mortality = child_mortality_0_5_year_olds_dying_per_1000_born, 
         pop = population_total) %>% 
  filter(between(time, 1945, 2018)) %>%
  ggplot(aes(child, (1000-mortality)/1000, size = pop)) +
  annotate(geom="rect", fill = NA, col = "gray40", 
           xmin = 8.5, xmax = 5, ymin = .55, ymax = .95) +
  annotate(geom="rect", fill = NA, col = "gray40", 
           xmin = 3.5, xmax = 1.1, ymin = .90, ymax = 1) +
  geom_point(shape = 21, fill = "gray40", alpha = .8) +
  scale_y_continuous("Children surviving to age 5",
                     breaks = 5:10/10,
                     limits = c(.5, 1),
                     labels = scales::percent) +
  scale_x_reverse("Babies per woman",
                  breaks = 1:8) +
  annotate(geom="text", col = "gray40", 
           x = 8.5, y = .55, vjust = 1, hjust = 0,
           label = "Big families and many children die") +
    annotate(geom="text", col = "gray40", 
           x = 3.5, y = .9, vjust = 1, hjust = 0,
           label = "Small families and\nfew children die") +
  theme(legend.position = "none") +
  labs(title = "Year: {frame_time}") +
  # Animation
  transition_time(time)
animate(p, nframes = 100)
```


# Acknowledgements

systemaGlobalis is simply a minimal R package wrapper to the dataset created by Gapminder. As such, please refer to <https://www.gapminder.org/> for more information about the data, including sources, and please acknowledge them for any reuse of the data.

>Gapminder created this dataset and provides it under Creative Common Attribution 4.0 International.

The R package is therefore distributed under the same license.

There is another gapminder R package that contains a selection of the gapminder data, [gapminder](https://github.com/jennybc/gapminder), which inspired this package.
